---
title: "Compositional trait MVMR"
output: html_notebook
---

```{r}
library(simulateGP)
library(tidyverse)
library(TwoSampleMR)
```

## Simulation questions
 
- What happens when you do straightforward GWAS of compositional phenotypes and feed into MR?
- What happens when there is measurement error?
  - e.g. 3 traits (buckets) - some of the signal in bucket 1 goes into each of the other two buckets
  - e.g. 3 traits (buckets) - some of the signal in bucket 1 goes into both of the other two buckets


## Compositional phenotypes

Three traits

- Active
- Sleeping
- Sedentary

### Scaling method

Randomly generate each phenotype independently, representing the proportion of time, and then scale so that they add up to 24 hours

```{r}
nsnp_active = 100
nsnp_sleeping = 100
nsnp_sedentary = 100
h2_active = 0.4
h2_sleeping = 0.4
h2_sedentary = 0.4

eff_active <- choose_effects(nsnp_active, h2_active)
eff_sleeping <- choose_effects(nsnp_sleeping, h2_sleeping)
eff_sedentary <- choose_effects(nsnp_sedentary, h2_sedentary)

geno_active <- make_geno(10000, nsnp_active, 0.5)
geno_sleeping <- make_geno(10000, nsnp_sleeping, 0.5)
geno_sedentary <- make_geno(10000, nsnp_sleeping, 0.5)

y_active <- make_phen(eff_active, geno_active, vy=1, my=8)
y_sleeping <- make_phen(eff_sleeping, geno_sleeping, vy=1, my=8)
y_sedentary <- make_phen(eff_sedentary, geno_sedentary, vy=1, my=8)

geno <- cbind(geno_active, geno_sleeping, geno_sedentary)
phen <- tibble(y_active=y_active, y_sleeping=y_sleeping, y_sedentary=y_sedentary)
dim(geno)
dim(phen)
rm(y_active, y_sleeping, y_sedentary)
```

Scale phenotypes

```{r}
phen_scale <- phen %>%
  mutate(s = y_active + y_sleeping + y_sedentary) %>%
  mutate(y_active = y_active / s * 24, y_sleeping=y_sleeping / s * 24, y_sedentary=y_sedentary / s * 24) %>%
  select(-c(s))
```

Outcome

```{r}
bxy = c(-0.1, -0.1, 0.1)
phen$chd <- make_phen(bxy, phen)
phen_scale$chd <- make_phen(bxy, phen_scale)
hist(phen$chd)
```

MVMR without scaling

```{r}
dat <- make_mvdat(list(active=phen$y_active, sedentary=phen$y_sedentary, sleeping=phen$y_sleeping), phen$chd, geno)
str(dat)
```
```{r}
mv_multiple(dat)
```

MVMR with scaling

```{r}
dat <- make_mvdat(list(active=phen_scale$y_active, sedentary=phen_scale$y_sedentary, sleeping=phen_scale$y_sleeping), phen_scale$chd, geno)

str(dat)
summary(lm(dat$outcome_beta ~ 0 + dat$exposure_beta, weight=1/dat$outcome_se^2))
```

Try removing the sleeping trait

```{r}
dat <- make_mvdat(list(active=phen_scale$y_active, sedentary=phen_scale$y_sedentary), phen_scale$chd, geno)
(z <- mv_multiple(dat))
```

Remove active

```{r}
dat <- make_mvdat(list(sleeping=phen_scale$y_sleeping, sedentary=phen_scale$y_sedentary), phen_scale$chd, geno)
(w <- mv_multiple(dat))
```

Remove sedentary

```{r}
dat <- make_mvdat(list(active=phen_scale$y_active, sleeping=phen_scale$y_sleeping), phen_scale$chd, geno)
(y <- mv_multiple(dat))
```

Try to find a solution for vector of effects that will return estimates for each?

```{r}
fn <- function(x, y1, y2, w2, w3, z1, z3)
{
  x_upda <- rep(0, 3)
  x_updb <- rep(0, 3)
  x_upda[1] <- y1 + x[3]
  x_updb[1] <- z1 + x[2]
  x_updb[2] <- w2 + x[1]
  x_upda[2] <- y2 + x[3]
  x_upda[3] <- w3 + x[1]
  x_updb[3] <- z3 + x[2]
  return(sum((x_upda-x_updb)^2))
}

# All the following give solutions, showing that the system is unidentified
optim(c(-0.1,-0.1,0.1), fn, 
      y1=y$result$b[1],
      y2=y$result$b[2],
      w2=w$result$b[2],
      w3=w$result$b[1],
      z1=z$result$b[1],
      z3=z$result$b[2]
    )

optim(c(1,1,1), fn, 
      y1=y$result$b[1],
      y2=y$result$b[2],
      w2=w$result$b[2],
      w3=w$result$b[1],
      z1=z$result$b[1],
      z3=z$result$b[2]
    )

optim(c(2,2,2), fn, 
      y1=y$result$b[1],
      y2=y$result$b[2],
      w2=w$result$b[2],
      w3=w$result$b[1],
      z1=z$result$b[1],
      z3=z$result$b[2]
    )
```

This is just unidentified.

Straightforward MR

```{r}
dat_active <- get_effs(phen_scale$y_active, phen_scale$chd, geno)
dat_sleeping <- get_effs(phen_scale$y_sleeping, phen_scale$chd, geno)
dat_sedentary <- get_effs(phen_scale$y_sedentary, phen_scale$chd, geno)
mr(dat_active, method="mr_ivw")
mr(dat_sleeping, method="mr_ivw")
mr(dat_sedentary, method="mr_ivw")
```

Interpretation:

1. The unscaled phenotype MV analysis is unbiased
2. The straightforward MR appears somewhat biased due to pleiotropy
3. The scaled MV analysis appears unbiased if we interpret the associations as being against the baseline of the mean value for the last trait
4. It looks like the MR results for 2 traits are correct if we add the true causal effect of the missing trait, but there are an infinite set of solutions because the system is unidentified?



### Partitioning method

Divide 24 hours into 3 segments, i.e. 2 breakpoints. Each breakpoint has a mean time, and each individual is given a deviation from the mean time.

TODO...





